Unifying quasars with clumpy wind models

### 2015-07-30 paper.tex
nearly there for first draft

%% declare any packages used
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{graphicx}
\usepackage{color}
\usepackage{pdfpages}
\usepackage{appendix}
\usepackage{subfigure}
\usepackage{url}
%\usepackage{epsfig}
%\usepackage[dvips]{color}
%\usepackage{aabib}

@@

\title
{
Modelling quasar outflows:
clumpy winds and broad emission lines
}

@@

\begin{abstract} 
% Broad absorption lines (BALs) in the ultraviolet 
% are seen in $\sim20\%$ of quasi-stellar objects (QSOs). 
% Blue-shifted broad absorption lines (BALs) are the most direct evidence of 
% accretion disc `winds' in such systems; mass loaded outflows
% emanating from the disc that may be driven by line forces or
% magnetic processes. 
Various unification schemes for
quasars and luminous active galactic nuclei (AGN) have proposed
that the broad emission line region is roughly cospatial
with broad absorption line (BAL) gas and much of the phenomenology of luminous AGN
can be explained by a simple geometrical picture involving an accretion
disc and associated outflow. Here, we test this paradigm by 
utilising our state-of-the-art radiative transfer code to produce synthetic spectra
from simple biconical disc wind models. In particular, we expand on our previous 
work in which a benchmark model for BAL quasars was produced. 
We find that a simple treatment of clumping (`microclumping') 
allows for a more realistic X-ray luminosity in the model by lowering the 
ionization parameter. We examine the X-ray properties of this new model
and find good agreement with existing X-ray samples of AGN and QSOs.
We find that clumping enhances the H recombination and collisionally excited resonance lines,
causing strong line emission (EW=?) to emerge at the low inclination angles, 
which represent quasars within this unification scenario.
% We also treat Hydrogen using a `macro-atom' approach in order to 
% examine the effect of recombination on Hydrogen emission lines, and
% this results in significant line emission in Lyman~$\alpha$ and the Balmer series.
However, we are unable to produce line emission with comparable equivalent widths
to existing quasar composites, due to a fundamental constraint arising
from the anisotropy of emission from a classical thin disc. We briefly 
explore the effect of relativistic beaming, gravitational redshift and 
light bending on the angular distribution of disc continuum emission. 
We find that these general relativistic effects do cause the disc to
emit more isotropically, but this is not yet sufficient to produce a
self-consistent model. We discuss a number of potential solutions. 
Overall, our work suggests that geometric unification
involving an accretion disc wind is a promising scenario, but our results 
pose a number of difficult challenges to such a model.
% Determining the true geometry of ADWs and uncovering the true disc spectral 
% (and angular) energy distribution are key next steps if we are to build up a 
% holistic picture of the quasar population.
\end{abstract}

@@

Approximately $20\%$ of quasars exhibit broad absorption lines (BALs) in the ultraviolet,
providing clear evidence for outflowing absorbing material
\citep{weymann1991, reichard2003, knigge2008, turnermiller2009, allen2011}.
The simplest explanation for the incidence of 
BAL quasars (BALQSOs) is in terms of an accretion disc wind (ADW). 
Within this paradigm, the BALQSO fraction is associated with
the covering factor of the outflow.
In addition, ADWs may offer a natural explanation for the
diverse phenemonology of luminous AGN and QSOs \citep[e.g.][]{MCGV95, elvis2000}. 
In such a model, a biconical wind rises from 
the accretion disc, and the class of object is explained by the material
intercepting the line of sight. Depending on viewing angle, an observer 
may then see a BALQSO or normal `Type 1' quasar.
Within this framework, the broad-line region (BLR) is normally
assumed to correspond to the dense wind base.
There are various spectroscopic sub-classifications of BALQSOs: 
HiBALQSOs, which only exhibit higher
ionization line absorption; LoBALQSOs which also show
absorption in lower ionization state species such as \mg\ and \al; and
FeLoBALQSOs which show further absorption in Fe~\textsc{ii} and \textsc{iii}.
In unified models, this is generally attributed to ionization stratification
of the outflow \citep[e.g.][]{elvis2000}.

% As well as acting as a source of photoionized plasma, a 
As well as imprinting clear line absorption features,
disc wind may also have a profound effect on the structure and 
emergent {\em continuum} of the accretion disc itself.
Mass-loss will alter the accretion rate and resultant 
temperature of the accretion disc, possibly explaining some 
of the features we typically see in luminous AGN \citep{laordavis2014}.
Recent results from \cite{capellupo2015} find 
that if one includes a combination of mass-loss, general relativity (GR) and Comptonisation
then AGN spectral energy distributions (SEDs) can, in general, be fitted well with accretion disc models.
Mass-loss therefore appears to be critical if an $\alpha$-disc
model is to successfully fit AGN SEDs, particular in the UV region of the spectrum.

@@

The efficiency of line-driving is crucially dependent on the ionization state 
of the outflowing plasma. \cite{MCGV95} proposed a potential solution: 
a region of `hitchhiking gas' that could shield the wind from the central X-ray source. 
Hydrodynamic simulations of line-driven disc winds also found a shielding region
was required to maintain the correct ionization state \citep{PSK2000,PK04}. 
However, \cite{H14} showed that including multiple scattering means the ionizing radiation 
field could still reach the previously shielded regions in those particular models.
An alternative solution is that the wind is clumped \citep[e.g.][]{hamann2013}
possibly on multiple scale lengths. Local density enhancements could lower the 
ionization parameter of the plasma while still maintaining the same mass-loss 
rate and column density.

@@

There has been some success using simple
kinematic prescriptions for biconical disc winds to model 
AGN and quasar outflows \citep{simlong2008,higginbottom2013}. 
\cite[][hereafter H13]{higginbottom2013},
successfully produced a benchmark model for BALQSOs. However, the model
had two key drawbacks. Firstly, an unrealistically low X-ray luminosity
was required in order to prevent over-ionization of the outflow.
Secondly, the model failed to produce the strong emission lines 
required at low inclinations in a unified model.
In this paper, we attempt to address these issues, and test the disc wind 
unification model using Monte Carlo radiative transfer (MCRT) and photoionization 
calculations. The paper is organised as follows.
In section 2, we describe our code. In section 3, we outline the model, including 
a description of our clumping prescription. In section 4, we present the results 
from a clumped model. In section 5 we discuss our results, 
focussing particular on the anisotropy of 
disc emission and GR effects, and finally, in section 6, we summarise our findings.

@@

We use the MCRT code \py\ to carry out our radiative transfer and photoionization
calculations in non-local thermodynamic equilibrium (non-LTE). 
The code is described extensively by a series of authors (LK02, S05, H13, M15).
For that reason we only briefly describe the key elements of the global 
ionization calculation and other important aspects.

@@

To treat line transfer, we adopt the same hybrid scheme 
described by M15, 
in which the energy flows
through the system are described in terms of indivisible
energy quanta of radiant or kinetic energy 
(`$r-$packets' and `$k-$packets' respectively; see also section 2.3).
These energy packets interacting with either `simple ions'
or `macro-atoms'. The macro-atom implementation 
is described in full by \cite{lucy2002, lucy2003}.
The scheme allows one to treat non-LTE line transfer in radiative equilibrium
without approximation for elements which are identified as 
full macro-atoms, while maintaining the fast `two-level' 
treatment of resonance lines when elements are identified 
as simple-ions. In this study,
only H \& He are treated as macro-atoms, because 
we expect recombination to be important
in determining their level populations (see M15).

@@

To avoid approximating the complex AGN SED as e.g. a dilute blackbody \citep[][M15]{ML93},
we model the SED in a cell using the technique described by H13. In this scheme,
the mean intensity in a series of bands is modeled as a normalised power law in 
frequency $\nu$
\begin{equation}
J_{\nu,j}=K_{pl}\nu^{\alpha_{pl}},
\end{equation}
for a band $j$, or an exponential 
\begin{equation}
J_{\nu,j}=K_{exp} e^{(-h\nu/k T_{exp})}.
\end{equation}
Here, $K_{exp}$, $K_{pl}$, $T_{exp}$ and $\alpha_{pl}$ are spectral fit parameters
deduced from the band-limited radiation field estimators.
The ionization rate out of ion $i$ can then be written as 
\begin{equation}
R_{i,i+1}(J)= 
\displaystyle{
n_i \left(C_{i} n_e + 
\sum_{band~j=0}^{n}~
{\int_{\nu_j}^{~\nu_j+1}{J_{\nu,j}\sigma_i(\nu)\nu^{-1}d\nu}}
\right),}
\end{equation}
where $\sigma_i$ is the photoionization cross-section and $C_{i}$
represents the collisional ionization coefficient.
The recombination rate {\em into} ion $i$ is simply given by 
\begin{equation}
R_{i+1,i}(T_e) = (\alpha^i_{RR} + \alpha^i_{DR} + \alpha^i_{CR}) n_{i+1} n_e
\end{equation}
Where each $\alpha$ here is the recombination rate coefficient into the ground state of ion $i$.
The subscripts denote radiative, dielectronic and 
collisional (three-body) recombination respectively.
We thus neglect recombination to and from excited states in the simple-ion calculation. 
For simple-ions, we use a dilute Boltzmann equation to calculate 
the population of level $k$ in ionic stage $j$,
\begin{equation}
\frac{n_{jk}}{n_j} = \frac{W g_k}{z_j(T_R)} \exp(-E_k/kT_R).
\end{equation}
Here $z_j$ is the partition function of ionic stage $j$,
$T_R$ is the effective radiation temperature,
$E_k$ is the energy difference between level $k$ and the ground state,
and $g_k$ is the statistical weight of level $k$.

@@

We include all free-free, bound-free and bound-bound heating
and cooling processes in the model. For radiative transfer purposes
we treat electron scattering in the Thomson limit, 
but take full account of Compton heating and cooling when
calculating the thermal balance of the plasma (see H13).
Adiabatic cooling is included and represents the only
departure from strict radiative equilibrium, but is insignificant 
in most of the outflow.
It is dealt with self consistently by the spontaneous destruction
of $k$-packets with a probability

\begin{equation}
P_{\rm{adiabatic~destruction}} = \frac{C_{\rm{adiabatic}}}{C_{\rm{total}}},
\end{equation}
where $C_{\rm{adiabatic}}$ and ${C_{\rm{total}}}$ represent the 
adiabatic and total cooling rates in the cell.

@@

\begin{figure} 
\centering
\includegraphics[width=0.45\textwidth]{figures/fig2_cartoon.eps}
\caption
{
A cartoon showing the geometry and some key parameters of
our biconical wind model.
}
\label{fig:cartoon}
\end{figure}

@@

The accretion disc in our model is geometrically thin, but optically thick
and this adopt a standard multi-temperature blackbody
using a \cite{shakurasunyaev1973} temperature profile. 
The emergent SED is thus determined by the specified accretion rate ($\dot{m}$)
and central BH mass ($M_{BH}$).
The inner radius of the disc extends to the innermost 
stable circular orbit (ISCO) of the BH. 
We assume a Schwarzchild BH with an ISCO at $6~R_G$.
The X-ray source in our model is treated as an isotropic sphere at the ISCO,
which emits photons according to a power law with index $\alpha_X$.
The normalisation of this power law is such that it produces the specified 
2-10kev luminosity.
In addition to the disc and X-ray source, 
the wind is able to reprocess radiation. However, new 
photon packets are not produced in the wind (as in LK02). 
Instead, this reprocessing is dealt with by enforcing strict
radiative equilibrium via an indivisible energy packet
constraint (see Lucy 2002, M15).
% \begin{equation}

@@

Describe briefly the kinematics and geometry of the model.

@@

Clumping the wind has an important effect on the ionization state and has
been proposed as a solution to the so-called `over-ionization problem' in 
disc winds (REFs). This is the main motivation for incorporating microclumping
into our model. This treatment is first-order; it does not adequately
represent the complex substructures and stratifications in ionization
state we expect in AGN outflows. Nevertheless, clumping is clearly
important in these flows, and this parameterization allows a simple estimate
for the effect clumping might have on the ionization state and emergent 
line emission. It is also encouraging that microclumping has been used 
successfully in fits to O-star wind spectra (REFs).

@@

Here we describe the results from our model. This set of parameters
was arrived at by conducting a limited grid search over a 
5-dimensional parameter space involving the variables
$r_{min}$, $\theta_{min}$, $f$, $\alpha$ and $R_V$.
We then evaluated these models qualitavely based on the following
criteria:

\begin{itemize}
\item Does the model maintain the correct ionization state and produce strong BALs?
\item Does strong line emission emerge at low inclinations?
\item Do H recombination lines appear in the spectrum?
\item Do a certain range of angles produce LoBAL features?
\end{itemize}

We arrived at a model which is one of the most promising,
but also representative of a family of models. Examination 
of this individual model allows us to gain insight
into fundamental geometrical and physical constraints, 
as discussed in section 5. 
The full grid, including output spectral files and plots can be found at
\url{jhmatthews.github.io/quasar-wind-grid/}.

\begin{figure*} %fullpage
\centering
\includegraphics[width=1.0\textwidth]{figures/wind.eps}
\caption
{
Physical properties of the outflow, shown by the coloured contours.
The solid black line marks a sphere at $1000~R_G$.
The dotted lines show the $75^\circ$ and $80^\circ$ sightlines 
to the centre of the system, and illustrate that different sightlines
intersect material of different ionization states.
}
\label{fig:uvspec}
\end{figure*} %fullpage

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{figures/sightlines.eps}
\caption
{
The classes of sightlines and the corresponding possible 
system types in the framework of our model.
%[FIGURE NEEDS IMPROVING]
}
\label{fig:lobal}
\end{figure}

\begin{table}
\begin{tabular}{p{3cm}p{4cm}}
\hline Free Parameters 	&	 Value \\ 
\hline \hline 
$M_{BH}$ 	 &	 $1\times 10^9~\rm{M_{\odot}}$ \\ 
$\dot{M}_{acc}$ 	 &	 $5~M_{\odot}yr^{-1} \simeq 0.2~\dot{M}_{Edd}$\\ 
$\alpha_X$ 	 &	 $-0.9$ \\ 
$L_{X} $ 	 &	 $10^{45}~\rm{ergs~s^{-1}}$\\ 
$r_{disc}(min)=r_{X}$   &	 $6r_g=8.8\times10^{14}~{\rm cm}$ \\ 
$r_{disc}(max)$   &	 $3400r_g = 5\times10^{17}~{\rm cm}$ \\ 
$\dot{M}_{wind}$  &	 $5~M_{\odot}yr^{-1}$ \\ 
$r_{min}$ 	&	 $300r_{g} = 4.4\times10^{16}~{\rm cm}$\\ 
$r_{max}$ 	&	 $600r_{g} = 8.8\times10^{16}~{\rm cm}$ \\ 
$\theta_{min}$ 	&	 $70.0^{\circ}$ \\ 
$\theta_{max}$ 	&	 $82.0^{\circ}$ \\ 
$\lambda$ 	&	 $0$ \\ 
$v_{\infty}$ 	&	 $v_{esc}$(f=1) \\ 
$R_v$ 	        &	 $10^{19}$cm \\ 
$\alpha$ 	&	 $1.0$ \\
$f$ 	&	 $0.01$ \\
\hline 
% Derived Parameters 	&	 Value \\ 
% \hline \hline
% $L_{\nu}(2500\mbox{\scriptsize{\AA}})$&	 $6.3\times10^{30}~\rm{ergs~s^{-1}~Hz^{-1}}$\\ 
% $L_{\nu}(2keV)$   &	 $1.2\times10^{25}~\rm{ergs~s^{-1}~Hz^{-1}}$\\ 
% $L_{bol}$ 	 &	 $2.4\times10^{46}~\rm{ergs~s^{-1}}$\\
% $M_{bol}$ 	 &	 -27.4\\ 
% $M_u$            &	 -26.2\\ 
%  $\alpha_{OX} $ 	 &	 -2.2\\ 
\end{tabular}
\caption{Wind geometry parameters used in the model.}
\label{wind_param}
\end{table}

@@

The wind rises slowly from the disc at first, with clumped densities
of $n_H \sim 10^{11}~\rm{cm^{-3}}$ close to the disc plane.
The flow then accelerates over a scale length of $R_V=10^{19}~\rm{cm}$
up to a terminal velocity of around $3$ times the escape velocity 
($\sim10,000~\rm{km~s^{-1}}$). This gradual acceleration means that
the wind exhibits a stratified ionization structure, with low ionization material
in the base of the wind giving way to highly ionized plasma further out.
By clumping the wind, we are able to produce the range of ionization states observed
in Quasars and BALQSOs, while adopting a realistic $2-10$ keV X-ray luminosity
of $L_{X}=10^{45}~\rm{ergs~s^{-1}}$. Without clumping, this wind would be over-ionized 
to the extent that opacities in e.g., \civ\ would be entirely negligible (see H13).

>>>>

%% declare any packages used
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{graphicx}
\usepackage{color}
\usepackage{pdfpages}
\usepackage{appendix}
\usepackage{subfigure}
\usepackage{url}
%\usepackage{pngfig}
%\usepackage[dvips]{color}
%\usepackage{aabib}

@@

\title
{
% Modelling quasar outflows:
% clumpy winds and broad emission lines
Unifying quasars with clumpy wind models
}

@@

\begin{abstract} 
% Broad absorption lines (BALs) in the ultraviolet 
% are seen in $\sim20\%$ of quasi-stellar objects (QSOs). 
% Blue-shifted broad absorption lines (BALs) are the most direct evidence of 
% accretion disc `winds' in such systems; mass loaded outflows
% emanating from the disc that may be driven by line forces or
% magnetic processes. 
Various unification schemes for
quasars and luminous active galactic nuclei (AGN) have proposed
that the broad emission line region is roughly cospatial
with broad absorption line (BAL) gas and much of the phenomenology of luminous AGN
can be explained by a simple geometrical picture involving an accretion
disc and associated outflow. Here, we test this paradigm by 
utilising our state-of-the-art radiative transfer code to produce synthetic spectra
from simple biconical disc wind models. In particular, we expand on our previous 
work in which a benchmark model for BAL quasars was produced. 
We have conducted a limited parameter search with the aim of unifying quasar phenomenology,
and arrive at an improved model. The grid is now publicly available.
We find that a simple treatment of clumping (`microclumping') 
allows for a more realistic X-ray luminosity in the model by lowering the 
ionization parameter. We examine the X-ray properties of this new model
and find good agreement with existing X-ray samples of AGN and QSOs.
We find that the dense, X-ray heated wind 
produces strong H recombination and collisionally excited resonance 
line emission to emerge at the low inclination angles, 
which represent quasars within this unification scenario.
% We also treat Hydrogen using a `macro-atom' approach in order to 
% examine the effect of recombination on Hydrogen emission lines, and
% this results in significant line emission in Lyman~$\alpha$ and the Balmer series.
However, we are unable to reproduce the remarkably similar of line-to-continuum
ratios between BAL and non-BAL quasar composites.
This is due to a fundamental constraint arising
from the anisotropy of emission from a classical thin disc. 
We briefly explore the effect of relativistic beaming, gravitational redshift and 
light bending on the angular distribution of disc continuum emission. 
We find that these general relativistic effects do cause the disc to
emit slightly more isotropically, but this is not yet sufficient to produce a
self-consistent model. We discuss wind reprocessing as a potential solution.
Overall, our work suggests that geometric unification
involving an accretion disc wind is a promising scenario, but our results 
pose a number of difficult challenges to such a model.
% Determining the true geometry of ADWs and uncovering the true disc spectral 
% (and angular) energy distribution are key next stpng if we are to build up a 
% holistic picture of the quasar population.
\end{abstract}

@@

Approximately $20\%$ of quasars exhibit blueshifted ($\sim 0.1c$)
broad absorption lines (BALs) in the ultraviolet,
providing clear evidence for outflowing absorbing material
\citep{weymann1991, reichard2003, knigge2008, turnermiller2009, allen2011}.
The simplest explanation for the incidence of 
BAL quasars (BALQSOs) is in terms of an accretion disc wind (ADW). 
Within this paradigm, the BALQSO fraction is associated with
the covering factor of the outflow.
In addition, ADWs may offer a natural explanation for the
diverse phenemonology of luminous AGN and QSOs \citep[e.g.][]{MCGV95, elvis2000}. 
In such a model, a biconical wind rises from 
the accretion disc, and the class of object is explained by the material
intercepting the line of sight. Depending on viewing angle, an observer 
may then see a BALQSO or normal `Type 1' quasar.
Within this framework, the broad-line region (BLR) is normally
assumed to correspond to the dense wind base.
There are various spectroscopic sub-classifications of BALQSOs: 
HiBALQSOs, which only exhibit higher
ionization line absorption; LoBALQSOs which also show
absorption in lower ionization state species such as \mg\ and \al; and
FeLoBALQSOs which show further absorption in Fe~\textsc{ii} and \textsc{iii}.
In unified models, this is generally attributed to ionization stratification
of the outflow \citep[e.g.][]{elvis2000}.

% As well as acting as a source of photoionized plasma, a 
As well as imprinting clear line absorption features,
disc winds may also have a profound effect on the structure and 
emergent {\em continuum} of the accretion disc itself.
Mass-loss will alter the accretion rate and resultant 
temperature of the accretion disc, possibly explaining some 
of the features we typically see in luminous AGN \citep{laordavis2014}.
Recent results from \cite{capellupo2015} find 
that if one includes a combination of mass-loss, general relativity (GR) and Comptonisation
then AGN spectral energy distributions (SEDs) can, in general, be fitted well with accretion disc models.
Mass-loss therefore appears to be critical if an $\alpha$-disc
model is to successfully fit AGN SEDs, particular in the UV region of the spectrum.

@@

The efficiency of line-driving is crucially dependent on the ionization state 
of the outflowing plasma, meaning that it is difficult to prevent 
the wind becoming over-ionized and `failing' in the presence of strong X-rays. 
\cite{MCGV95} proposed a potential solution: 
a region of `hitchhiking gas' that could shield the wind from the central X-ray source. 
Hydrodynamic simulations of line-driven disc winds also found a shielding region
was required to maintain the correct ionization state \citep{PSK2000,PK04}. 
However, \cite{H14} showed that including multiple scattering means the ionizing radiation 
field could still reach the previously shielded regions in those particular models.
An alternative solution is that the wind is clumped \citep[e.g.][]{hamann2013}
possibly on multiple scale lengths. Local density enhancements could lower the 
ionization parameter of the plasma while still maintaining the same mass-loss 
rate and column density.

@@

There has been some success using simple
kinematic prescriptions for biconical disc winds to model 
AGN and quasar outflows \citep[][hereafter H13]{simlong2008,sim2010,higginbottom2013}. 
H13 successfully produced a benchmark model for BALQSOs. However, the model
had two key drawbacks. Firstly, an unrealistically low X-ray luminosity
was required in order to prevent over-ionization of the outflow.
Secondly, the model failed to produce the strong emission lines 
required at low inclinations in a unified model.
In this paper, we attempt to address these issues, and test the disc wind 
unification model using Monte Carlo radiative transfer (MCRT) and photoionization 
calculations. The paper is organised as follows.
In section 2, we describe some of the important photoionization 
and MCRT aspects of the code. In section 3, we outline the model, including 
a description of our clumping implementation. In section 4, we present the results 
from a clumped model. In section 5 we discuss our results, 
focussing particular on the anisotropy of 
disc emission and GR effects, and finally, in section 6, we summarise our findings.

@@

We use the MCRT code \py\ to carry out our radiative transfer and photoionization
calculations in non-local thermodynamic equilibrium (non-LTE). 
The has been used to model accreting white dwarfs 
(Long \& Knigge 2002, hereafter LK02; Noebauer et al. 2010; 
Matthews et al. 2015, hereafter M15), young-stellar objects \citep{simmacro2005}
and quasars (H13). \nocite{noebauer, M15, LK02}
These studies contain extensive detail on the code, 
so we only briefly describe the key elements of the global 
ionization calculation and other important aspects.

@@

To treat line transfer, we adopt the same hybrid scheme 
described by M15, 
in which the energy flows
through the system are described in terms of indivisible
energy quanta of radiant or kinetic energy 
(`$r-$packets' and `$k-$packets' respectively; see also section 2.3).
These energy packets interact with either two-level `simple ions'
or full `macro-atoms'. The macro-atom implementation 
is described in full by \cite{lucy2002, lucy2003}.
Our scheme allows one to treat non-LTE line transfer in radiative equilibrium
without approximation for elements which are identified as 
full macro-atoms, while maintaining the fast `two-level' 
treatment of resonance lines when elements are identified 
as simple-ions (see M15). In this study,
only H \& He are treated as macro-atoms, because 
we expect recombination to be important
in determining their level populations and resultant line emission.

@@

To avoid approximating the complex AGN SED as e.g. a dilute blackbody \citep[][M15]{ML93},
we model the SED in a cell using the technique described by H13. In this scheme,
the mean intensity in a series of bands is modeled as a normalised power law in 
frequency $\nu$
\begin{equation}
J_{\nu,j}=K_{pl}\nu^{\alpha_{pl}},
\end{equation}
for a band $j$, or an exponential 
\begin{equation}
J_{\nu,j}=K_{exp} e^{(-h\nu/k T_{exp})}.
\end{equation}
Here, $K_{exp}$, $K_{pl}$, $T_{exp}$ and $\alpha_{pl}$ are spectral fit parameters
deduced from the band-limited radiation field estimators.
The ionization rate out of ion $i$ can then be written as 
\begin{equation}
R_{i,i+1}(J)= 
\displaystyle{
n_i \left(C_{i} n_e + 
\sum_{band~j=0}^{n}~
{\int_{\nu_j}^{~\nu_j+1}{J_{\nu,j}\sigma_i(\nu)\nu^{-1}d\nu}}
\right),}
\end{equation}
where $\sigma_i$ is the photoionization cross-section and $C_{i}$
represents the collisional ionization coefficient.
The recombination rate {\em into} ion $i$ is simply given by 
\begin{equation}
R_{i+1,i}(T_e) = (\alpha^i_{RR} + \alpha^i_{DR} + \alpha^i_{CR}) n_{i+1} n_e
\end{equation}
Where each $\alpha^i$ here is the recombination rate coefficient into the ground state of ion $i$.
The subscripts denote radiative, dielectronic and 
collisional (three-body) recombination respectively.
We thus neglect recombination to and from excited states in the simple-ion calculation. 
For simple-ions, we use a dilute Boltzmann equation to calculate 
the population of level $k$ in ionic stage $j$,
\begin{equation}
\frac{n_{jk}}{n_j} = \frac{W g_k}{z_j(T_R)} \exp(-E_k/kT_R).
\end{equation}
Here $z_j$ is the partition function of ionic stage $j$,
$T_R$ is the effective radiation temperature,
$E_k$ is the energy difference between level $k$ and the ground state,
and $g_k$ is the statistical weight of level $k$.

@@

We include all free-free, bound-free and bound-bound heating
and cooling processes in the model. For radiative transfer purposes
we treat electron scattering in the Thomson limit, 
but take full account of Compton heating and cooling when
calculating the thermal balance of the plasma (see H13).
Adiabatic cooling is included and represents the only
departure from strict radiative equilibrium, but is insignificant 
in most of the outflow.
% It is dealt with self consistently by the spontaneous destruction
% of $k$-packets with a probability

% \begin{equation}
% P_{\rm{adiabatic~destruction}} = \frac{C_{\rm{adiabatic}}}{C_{\rm{total}}},
% \end{equation}
% where $C_{\rm{adiabatic}}$ and ${C_{\rm{total}}}$ represent the 
% adiabatic and total cooling rates in the cell.

@@

\begin{figure} 
\centering
\includegraphics[width=0.45\textwidth]{figures/fig2_cartoon.png}
\caption
{
A cartoon showing the geometry and some key parameters of
our biconical wind model.
}
\label{fig:cartoon}
\end{figure}

@@

The accretion disc in our model is geometrically thin, but optically thick
and thus adopt a standard multi-temperature blackbody
using a \cite{shakurasunyaev1973} temperature profile.

The emergent SED is thus determined by the specified accretion rate ($\dot{m}$)
and central BH mass ($M_{BH}$).
All photon sources in our model are assumed opaque, meaning
that photons which strike them are destroyed.
The inner radius of the disc extends to the innermost 
stable circular orbit (ISCO) of the BH. 
We assume a Schwarzchild BH with an ISCO at $6~R_G$.
The X-ray source in our model is treated as an isotropic sphere at the ISCO,
which emits r-packets according to a power law in flux with index $\alpha_X$ such that
\begin{equation}
F_X (\nu) = K \nu^{\alpha_X}.
\end{equation}
The normalisation, $K_X$ of this power law is such that it 
produces the specified 2-10~keV luminosity, $L_X$.
In addition to the disc and X-ray source, 
the wind is able to reprocess radiation. However, new 
photon packets are not produced in the wind (as in LK02). 
Instead, this reprocessing is dealt with by enforcing strict
radiative equilibrium ({\em modulo} adiabatic cooling)
via an indivisible energy packet
constraint (see Lucy 2002, M15; also section 2.3).
% \begin{equation}

@@

In this prescription, a biconical disc wind rises from the accretion 
disc between launch radii $r_{min}$ and $r_{max}$.
The opening angles of the wind are set to $\theta_{min}$ and $\theta_{max}$.
The poloidal velocity along each individual streamline at a poloidal distance $l$ 
is then given by
\begin{equation}
v_l=v_0+\left[v_{\infty}(r_0)-v_0\right]\frac{\left(l/R_v\right)^{\alpha}}{\left(l/R_v\right)^{\alpha}+1},
\label{v_law}
\end{equation}
where $v_0$ is the velocity at the base of the streamline, $\alpha$ is
an exponent governing how quickly the wind accelerations and 
$R_v$ is the `acceleration length', defined as the distance at which
the outflow reaches half of its terminal velocity, $v_{\infty}$.
The terminal velocity is set to a fixed multiple of the escape
velocity, $v_{esc}$ at the base of the streamline (radius $r_0$).
The rotational velocity, $v_{\phi}$, is initially Keplerian ($v_k = [GM/r_0]^{1/2}$),
and the wind conserves specific angular momentum such that 
\begin{equation}
v_{\phi} r = v_k r_0.
\label{v_law}
\end{equation}
The velocity law is crucial in determining the output spectra,
as it affects not only the projected velocities along the line of sight,
but also the density and ionization state of the outflow.
A wind which accelerates more slowly will have a denser wind base
with correspondingly different ionization and emission characteristics.

@@

Clumping the wind has an important effect on the ionization state and has
been proposed as a solution to the so-called `over-ionization problem' in 
disc winds (REFs). This is the main motivation for incorporating microclumping
into our model. This treatment is first-order; it does not adequately
represent the complex substructures and stratifications in ionization
state we expect in AGN outflows. Nevertheless, clumping is clearly
important in these flows, and this parameterization allows a simple estimate
for the effect clumping might have on the ionization state and emergent 
line emission. It is also encouraging that microclumping has been used 
successfully in fits to O-star wind spectra \citep{hillier1991eswingsmodel}.

@@

Here we describe the results from our model, the parameters
of which are shown in table~1.
Parameters differing from the benchmark model of
H13 are highlighted with an asterisk.
This set of parameters
was arrived at by conducting a limited grid search over a 
5-dimensional parameter space involving the variables
$r_{min}$, $\theta_{min}$, $f$, $\alpha$ and $R_V$.
The full grid, including output spectral files and plots can be found at
\url{jhmatthews.github.io/quasar-wind-grid/}.
We evaluated these models qualitavely based on the following
criteria:
\begin{itemize}
\item Does the model maintain the correct ionization state and produce strong BALs?
\item Does strong line emission emerge at low inclinations?
\item Do H recombination lines appear in the spectrum?
\item Do a certain range of angles produce LoBAL features?
\item Does the model compare favourably to quasar composite spectra?
\end{itemize}
We arrived at a model which is one of the most promising,
but also broadly representative of a family of models. 
Examination of this individual model allows us to gain insight
into fundamental geometrical and physical constraints, 
as discussed in section 5, and assess the models potential
for unification.

\begin{figure*} %fullpage
\centering
\includegraphics[width=1.0\textwidth]{figures/wind.png}
\caption
{
Physical properties of the outflow, shown by the coloured contours.
The solid black line marks a sphere at $1000~R_G$.
The dotted lines show the $72^\circ$ and $78^\circ$ sightlines 
to the centre of the system, and illustrate that different sightlines
intersect material of different ionization states.
}
\label{fig:uvspec}
\end{figure*} %fullpage

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{figures/sightlines.png}
\caption
{
The classes of sightlines and the corresponding possible 
system types in the framework of our model.
%[FIGURE NEEDS IMPROVING]
}
\label{fig:lobal}
\end{figure}

\begin{table}
\begin{tabular}{p{3cm}p{4cm}}
\hline Free Parameters 	&	 Value \\ 
\hline \hline 
$M_{BH}$ 	 &	 $1\times 10^9~\rm{M_{\odot}}$ \\ 
$\dot{M}_{acc}$ 	 &	 $5~M_{\odot}yr^{-1} \simeq 0.2~\dot{M}_{Edd}$\\ 
$\alpha_X$ 	 &	 $-0.9$ \\ 
$L_{X} $ 	 &	 $10^{45}~\rm{ergs~s^{-1}}$$^*$ \\ 
$r_{disc}(min)=r_{X}$   &	 $6r_g=8.8\times10^{14}~{\rm cm}$ \\ 
$r_{disc}(max)$   &	 $3400r_g = 5\times10^{17}~{\rm cm}$ \\ 
$\dot{M}_{wind}$  &	 $5~M_{\odot}yr^{-1}$ \\ 
$r_{min}$ 	&	 $300r_{g} = 4.4\times10^{16}~{\rm cm}$\\ 
$r_{max}$ 	&	 $600r_{g} = 8.8\times10^{16}~{\rm cm}$ \\ 
$\theta_{min}$ 	&	 $70.0^{\circ}$ \\ 
$\theta_{max}$ 	&	 $82.0^{\circ}$ \\ 
$\lambda$ 	&	 $0$ \\ 
$v_{\infty}(r_0)$ 	&	 $v_{esc}(r_0)$ \\ 
$R_v$  	        &	 $10^{19}$cm$^*$ \\ 
$\alpha$ 	&	 $0.6^*$ \\
$f$ 	&	 $0.01^*$  \\
\hline 
% Derived Parameters 	&	 Value \\ 
% \hline \hline
% $L_{\nu}(2500\mbox{\scriptsize{\AA}})$&	 $6.3\times10^{30}~\rm{ergs~s^{-1}~Hz^{-1}}$\\ 
% $L_{\nu}(2keV)$   &	 $1.2\times10^{25}~\rm{ergs~s^{-1}~Hz^{-1}}$\\ 
% $L_{bol}$ 	 &	 $2.4\times10^{46}~\rm{ergs~s^{-1}}$\\
% $M_{bol}$ 	 &	 -27.4\\ 
% $M_u$            &	 -26.2\\ 
%  $\alpha_{OX} $ 	 &	 -2.2\\ 
\end{tabular}
\caption{Wind geometry parameters used in the model.}
\label{wind_param}
\end{table}

@@

The wind rises slowly from the disc at first, with clumped densities
of $n_H \sim 10^{11}~\rm{cm^{-3}}$ close to the disc plane.
The flow then accelerates over a scale length of $R_V=10^{19}~\rm{cm}$
up to a terminal velocity of around $3$ times the escape velocity 
($\sim10,000~\rm{km~s^{-1}}$). This gradual acceleration means that
the wind exhibits a stratified ionization structure, with low ionization material
in the base of the wind giving way to highly ionized plasma further out.
By clumping the wind, we are able to produce the range of ionization states observed
in wuasars and BALQSOs, while adopting a realistic $2-10$ keV X-ray luminosity
of $L_{X}=10^{45}~\rm{ergs~s^{-1}}$. Without clumping, this wind would be over-ionized 
to the extent that opacities in e.g., \civ\ would be entirely negligible (see H13).
